Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-11 paper.tex
moved attacks subsection to improve vertical distribution - fig. are not perfectly placed, though :(
merged paper.tex with submitted_version.tex (missing: replacement of Corona-virus warning app with CWA)

\section{Introduction}
\label{sec:intro}
The German Chancellor Angela Merkel has strongly supported the publication of
the mobile phone Corona warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
in Germany have downloaded the app with 6 million on the first day.
CWA is one amongst many similar applications that aim at the very important goal
to ``break infection chains'' by providing timely information to users of whether they
have been in close proximity to a person who tested positive for COVID-19.

%The Corona-virus warning app
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
%heated debate with supporters of a central architecture. The distributed architecture is
that is
based on a very clever application design whereby clients broadcast
highly anonymized so called ``Ephemeral IDs'' via Bluetooth and
store those ids of people in close proximity.
%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
%Bluetooth Low Energy Beacon protocol.
%The app saves those ids of people in close proximity.
Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
hence, stored contact data has never to be shared.
%Infected persons report their infection to a central server,
%which, in turn, provides all clients the means to reconstruct the ids
%used by reported cases over the last 14 days.
%
%When at a later date an infected person reports his infection to a central server,
%all clients are provided the means to reconstruct the ids used by this particular
%person's device over the last 14 days.
%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%The Corona-virus app has been produced based on a sophisticated security
%concept conceived by experts in the field %to our knowledge, no formal verification
%%
%and has strong claims with regard to mathematical support:
%
%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
%
However, there has been, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or to what extend trade offs have to be made due to inherent vulnerabilities.
Given that the protocol implemented by CWA resembles only the most basic
of the three protocols proposed by the DP-3T project,
a formal verified model that stresses the impact or limits of certain security measures
might give more weight to appeals like \cite{dp3t:impl}
%to Apple and Google
to adopt the more sophisticated protocols.
%
%Thereby, we believe
%that our current work is useful to increase the trust in the Corona-virus app necessary for
%its wide adoption which in turn is crucial for it to be efficient.

@@

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of the Corona-virus warning app (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the Corona-virus warning app (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

@@

Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
\begin{figure}
\vspace{-.5cm}
  \begin{center}
    \includegraphics[width=4in]{DP-3T_basic_protocol}
  \end{center}
\vspace{-.5cm}
  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
  \label{fig:dp3tprot}
\vspace{-.5cm}
\end{figure}

\subsubsection{Attack scenarios}
There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
The debate among %emerging between
advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
%As first step, both attacks start with deanonymizing Ephemeral Ids,
%i.\,e. linking them the particular devices.
Both attacks rely on deanonymisation of Ephemeral Ids,
i.\,e. linking ids to particular devices.

@@

\section{Summary and discussion of relevance of the approach}
%{Conclusions}
\label{sec:concl}
%\subsection{Protection goals of attacks}
We can establish in our formal framework an attack that even a system using changing
Ephemeral ids can be broken if the attacker physically follows a victim. This is a basic attack on
anonymity: a user's connection between his Ephemeral Ids and personal details (Iphone MAC or name)
is revealed to the attacker. The protection goal of privacy is thereby destroyed.

>>>>

\section{Introduction}
\label{sec:intro}
The German Chancellor Angela Merkel has strongly supported the publication of
the mobile phone Corona-virus warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
in Germany have downloaded the app with 6 million on the first day.
CWA is one amongst many similar applications that aim at the very important goal
to ``break infection chains'' by providing timely information to users of whether they
have been in close proximity to a person who tested positive for COVID-19.

%The CWA
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
%heated debate with supporters of a central architecture. The distributed architecture is
that is
based on a very clever application design whereby clients broadcast
highly anonymized so called ``Ephemeral IDs'' via Bluetooth and
store those ids of people in close proximity.
%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
%Bluetooth Low Energy Beacon protocol.
%The app saves those ids of people in close proximity.
Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
hence, stored contact data has never to be shared.
%Infected persons report their infection to a central server,
%which, in turn, provides all clients the means to reconstruct the ids
%used by reported cases over the last 14 days.
%
%When at a later date an infected person reports his infection to a central server,
%all clients are provided the means to reconstruct the ids used by this particular
%person's device over the last 14 days.
%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%The Corona-virus app has been produced based on a sophisticated security
%concept conceived by experts in the field %to our knowledge, no formal verification
%%
%and has strong claims with regard to mathematical support:
%
%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
%
However, there has, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or to what extent trade offs have to be made due to inherent vulnerabilities.
The protocol implemented by the framework CWA resembles only the most basic
of the three protocols proposed by the DP-3T project.
Hence, a formal verified model that stresses the impact or limits of certain security measures
might give more weight to appeals like \cite{dp3t:impl}
%to Apple and Google
to adopt the more sophisticated protocols.
%
%Thereby, we believe
%that our current work is useful to increase the trust in the Corona-virus app necessary for
%its wide adoption which in turn is crucial for it to be efficient.

@@

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of the German app (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the Corona-virus warning app (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
The debate among %emerging between
advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
%As first step, both attacks start with deanonymizing Ephemeral Ids,
%i.\,e. linking them the particular devices.
%Both attacks rely on deanonymisation of Ephemeral Ids,
%i.\,e. linking ids to particular devices.

@@

Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
\begin{figure}[htb]
\vspace{-.5cm}
  \begin{center}
    \includegraphics[width=4in]{DP-3T_basic_protocol}
  \end{center}
\vspace{-.5cm}
  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
  \label{fig:dp3tprot}
\vspace{-.5cm}
\end{figure}

%\subsubsection{Attack scenarios}
%There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
%The debate among %emerging between
%advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

%In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
%%As first step, both attacks start with deanonymizing Ephemeral Ids,
%%i.\,e. linking them the particular devices.
%Both attacks rely on deanonymisation of Ephemeral Ids,
%i.\,e. linking ids to particular devices.

@@

\section{Summary and discussion of relevance of the approach}
%{Conclusions}
\label{sec:concl}
%\subsection{Protection goals of attacks}
%\subsection{Summary and discussion of relevance of the approach}
We can establish in our formal framework an attack that even a system using changing
Ephemeral ids can be broken if the attacker physically follows a victim. This is a basic attack on
anonymity: a user's connection between his Ephemeral Ids and personal details (Iphone MAC or name)
is revealed to the attacker. The protection goal of privacy is thereby destroyed.
